---
title: "assignement1_eco"
format: html
editor: visual
---

## Assignment 1: Data and Descriptive

This assignment is intended to gather, process, and analyze sub-national GDP and population data for a selection of European countries.
We will calculate GDP per capita and investigate regional disparities using various descriptive statistics and visualizations.

We have been tasked with analyzing Ireland, Poland, Spain, Slovenia, the Netherlands, Greece, Norway, Albania from the year 2000 to 2020.

```{r}
#| setup: lable
library(tidyverse)
```

## Del A : Sub-national GDP and GDP per Capita

### **1 . Data Acquisition:**

#### Dataset: GDP

This dataset is a sub-regional GDP dataset for a selection of our European countries.
GDP stands for Gross Domestic Product, and it is a key economic indicator used to measure the overall economic health and performance of a country or region.
The variables relevant to us are;

GEO: this variable show us the geopolitical entity.

Time_Period: this variable shows us the time period in years.

Obs_Value: this variable shows us the value of GDP.

```{r}
library(tidyverse)
GDP <- read_csv("~/DataScience/GDP.csv")
View(GDP)
```

Before we start calculating the GDP per inhabitant for each sub-region, we want to "clean up" the dataset so that we only include the necessary variables:

```{r}
# We rename one of the variables and select the ones we want
GDP1 <- GDP |> 
  select(year = TIME_PERIOD, geo, OBS_VALUE)
```

#### Dataset: Population

This dataset is a dataset about population for a selection of our European countries.
Population refers to the total number of individuals of a particular species or group inhabiting a specific area or planet at a given time.The variables relevant to us are;

GEO: this variable show us the geopolitical entity.

Time_Period: this variable shows us the time period in years.

Obs_Value: this variable shows us the value of GDP.

```{r}
library(tidyverse)
population <- read_csv("~/DataScience/population.csv")
View(population)
```

we will do the same as on the previous data set by selecting the necessary variables:

```{r}
# We rename one of the variables and select the ones we want
population1 <- population |> 
  select(year = TIME_PERIOD, geo, OBS_VALUE)
```

### **2. GDP per Capita Calculation:** 
