---
title: "assignement1_eco"
format: html
editor: visual
---

## Assignment 1: Data and Descriptive

This assignment is intended to gather, process, and analyze sub-national GDP and population data for a selection of European countries.
We will calculate GDP per capita and investigate regional disparities using various descriptive statistics and visualizations.

We have been tasked with analyzing Ireland, Poland, Spain, Slovenia, the Netherlands, Greece, Norway, Albania from the year 2000 to 2020.

```{r}
#| setup: lable
library(tidyverse)
```

## Dataset: GDP

This dataset is a sub-regional GDP dataset for a selection of our European countries.
GDP stands for Gross Domestic Product, and it is a key economic indicator used to measure the overall economic health and performance of a country or region.
The variables relevant to us are;

GEO: this variable show us the geopolitical entity.

Time_Period: this variable shows us the time period in years.

Obs_Value: this variable shows us the value of GDP.

```{r}
library(tidyverse)
GDP <- read_csv("~/DataScience/GDP.csv")
View(GDP)
```

## Dataset: Population

This dataset is a dataset about population for a selection of our European countries.
Population refers to the total number of individuals of a particular species or group inhabiting a specific area or planet at a given time.The variables relevant to us are;

GEO: this variable show us the geopolitical entity.

Time_Period: this variable shows us the time period in years.

Obs_Value: this variable shows us the value of GDP.

```{r}
library(tidyverse)
population <- read_csv("~/DataScience/population.csv")
View(population)
```
